Unsupervised Detection and Tracking of Multiple Objects with Dependent Dirichlet Process Mixtures

### 2011-12-24 ltx/paper.tex
new edits. better svjour3 formatting. first draft of abstract. more author details.

\documentclass[smallcondensed, final]{svjour3}
% \documentclass[twocolumn, final]{svjour3}

@@

\title{Unsupervised Detection and Tracking of Multiple Objects with Dependent Dirichlet Process Mixtures}
\author{Willie Neiswanger and Frank Wood}
\maketitle
\mbox{}

\abstract{Main points.}

@@

Research in the past ten years involving the development of methods for general detection and tracking of objects has tended to focus on one of either extraction, localization, or tracking; attempts to integrate all three tasks and produce a system aimed at multiple arbitrary objects and diverse video types was not often a primary focus, or was attempted in an ad-hoc manner to test or apply a more specific technique. Research aimed at developing methods to accomplish the three tasks in a cohesive manner has been pursued more recently to produce algorithms that can, in a fully automated way, accomplish unsupervised detection and tracking--or in short, that can determine the spatiotemporal region of a video occupied by each distinct object. A widely applied, though unsophisticated, example of general detection and tracking is the technique of blob tracking. Blob tracking typically refers to methods that perform extraction and localization of objects at each frame independently--for example, by assuming that pixels representing an object are distinguishable from pixels representing the background via color, motion, or texture information, and segmenting the extracted data at each frame into ``blobs'' to accomplish localization. A number of tracking techniques based on finding blobs in consecutive frames of a video with similar locations and appearances have been attempted to maintain the identity of detected blobs over time. The main challenges, however, associated with blob tracking involve the difficulty of segmenting neighboring or occluding objects in a given frame \cite{zhao2004tracking} and preventing multiple tracked blobs from collapsing onto and tracking a single object \cite{vermaak_2003}. These difficulties stem from the typical blob tracker's need to perform detection via segmentation on individual frames and lack of a principled way to perform tracking while resisting the collapse of independent tracks.

This paper presents the use of a Bayesian nonparametric mixture model for general, unsupervised multiple object detection and tracking. This method is intended to provide a widely applicable way to detect arbitrary objects in videos--particularly in cases where frame-by-frame segmentation is difficult or video quality is low and extraction is noisy--and better maintain the isolation of independent objects during tracking. Specifically, this work describes how a Dirichlet process mixture model can accomplish these goals when it is applied to data that may be extracted in a general manner from videos. We will demonstrate how to extract data from objects in videos and how this generative model allows for automated detection and tracking. The model is first presented in an abstract form, to allow it to be used with arbitrary extraction procedures and object appearance representations that may be desired in future studies. We then formulate a specific model for data gained during a basic extraction procedure and describe inference algorithms that allow object localization and tracking to be carried out. We demonstrate our implementation on a number of synthetic and benchmark datasets and describe and compute performance metrics to quantify results.

@@

Methods for tracking single, non-rigid, objects are numerous. Color, for example, has long been used for tracking. These techniques often operate by modeling an object with some color-based appearance model, and using this model to find the object in subsequent frames. One classic approach involves modeling the color of a target (in particular, a distribution over the hue-saturation space of a target region) with a Gaussian mixture model (GMM), and choosing subsequent target positions by searching surrounding areas for regions that yield similar GMMs. Furthermore, the GMM is often allowed to adapt over time, and slowly change to model changes in lighting or other smooth time-based variations in the target's color \citep{raja_1998, mckenna_1999, jepson_2003}. Additionally, color has been successfully applied as the feature in a technique known a the mean-shift procedure (or the kernel-based method) for tracking \citep{comaniciu_2003, perez_2002, nummiaro_2003, lee_2011}. This procedure--a derivation and overview of which can be found in \citep{fukunaga_1975, cheng_1995}--optimizes the search for the region in a frame which is most similar to a target region in the previous frame. It involves an iterative algorithm that repeatedly shifts each data point to the weighted average of data points in its neighborhood; it has been proven that this process converges for each data point. Additionally, the process (in particular, the specification of the neighborhood and the weight-distribution when calculating the weighted mean of nearby data points) is generalized so that multiple kernels can be specified and used to allow for varied clustering behaviors. It can be shown that, through this procedure, each data point becomes associated with a local point of high density (dependent upon the underlying weight distribution specified by the kernel) which naturally allows for clustering \citep{cheng_1995}. The mean shift algorithm has been implemented successfully to allow for a sort of kernel-based object tracking \citep{comaniciu_2003, comaniciu_1999, comaniciu_2000}.

@@

% The mean-shift procedure (also known as `kernel-based object tracking') attempts to provide a robust way for non-rigid objects to be tracked. This method is beneficial because it optimizes the search for the `next' position of an object (i.e. the search for the region in a frame which is most similar to a target region in the previous frame). This procedure--a derivation and overview of which can be found in \citep{fukunaga_1975, cheng_1995}--involves an iterative algorithm that repeatedly shifts each data point to the weighted average of data points in its neighborhood; it has been proven that this process converges for each data point. Additionally, the process (in particular, the specification of the neighborhood and the weight-distribution when calculating the weighted mean of nearby data points) is generalized so that multiple kernels can be specified and used to allow for varied clustering behaviors. It can be shown that, through this procedure, each data point becomes associated with a local point of high density (dependent upon the underlying weight distribution specified by the kernel) which naturally allows for clustering \citep{cheng_1995}. The mean shift algorithm has been implemented successfully to allow for a sort of kernel-based object tracking \citep{comaniciu_2003, comaniciu_1999, comaniciu_2000}. Given the current position of an object at a given frame, the goal of tracking is often to find a nearby position in the next frame that has the most similar distribution over some common set of features. Usually this must be done through an exhaustive search, comparing the similarity of distributions at each nearby position with that at the current position. However, if a certain type of `isotropic kernel' (mean-shift kernel) known as the Bhattacharyya coefficient is chosen as a similarity metric between the feature distributions at two positions, it creates a smooth function where gradient descent techniques can be used to quickly converge upon an optimal subsequent position without using an exhaustive search. Many papers are concerned with applying this kernel-based object tracking scheme to different sets of features or with different kernels. Additionally, the scheme has been attempted with an adaptive kernel whose shape, scale, and orientation is influenced by a target being tracked \citep{wang_2004}, with a kernel adjusted by the estimated centroid of a tracked target \citep{mehmood_2009}, and with a heirarchical version of the mean-shift procedure \citep{dementhon_2002}.

@@

Recall that the distribution over the cluster assignments $c_{i}$ (one for each observation $i$) is a function of the cluster sizes (see \eqref{crp_rep}). Hence, to account for varying numbers of clusters--and in particular, to allow clusters to diminish in size (i.e. reduce the number of assigned observations) and even die--there is a deletion procedure (described in \cite{caron_2007}) by which observations are considered ``removed'' from their assigned clusters at a given time step and value of $m_{k}$ (defined in \eqref{crp_rep} to be the size of cluster $k$) is reduced.

@@

\begin{figure}
  \centering
  \subfloat[]{\label{fig:ants_img}\includegraphics[width=0.335\textwidth]{../img/usbh_frame_723.pdf}} 
  \subfloat[]{\label{fig:ants_img2}\includegraphics[width=0.335\textwidth]{../img/usbh_frame_724.pdf}}
  \subfloat[]{\label{fig:ants_img_framediff}\includegraphics[width=0.323\textwidth]{../img/usbh_framediff.png}}\\
  \subfloat[]{\label{fig:traffic2_img}\includegraphics[width=0.328\textwidth]{../img/traffic2_frame1.png}} 
  \subfloat[]{\label{fig:traffic2_img2}\includegraphics[width=0.328\textwidth]{../img/traffic2_frame2.png}}
  \subfloat[]{\label{fig:traffic2_img_framediff}\includegraphics[width=0.3273\textwidth]{../img/traffic2_framediff.png}}\\
  \subfloat[]{\label{fig:pets2009_img}\includegraphics[width=0.333\textwidth]{../img/pets2009_frame1.png}} 
  \subfloat[]{\label{fig:pets2009_img2}\includegraphics[width=0.333\textwidth]{../img/pets2009_frame2.png}}
  \subfloat[]{\label{fig:pets2009_img_framediff}\includegraphics[width=0.333\textwidth]{../img/pets2009_framediff.png}}\\
  \subfloat[]{\label{fig:longimg}\includegraphics[width=1.05\textwidth]{../img/longimg_6.pdf}} 
  \caption{Pairs of consecutive frames and the results produced by taking the pixel-wise difference between these frames. \willie{Want to add colorbar that explains how color represents frame number red = frame 1, blue = frame 100}}
  \label{fig:img_and_framediff}
\end{figure}

@@

The distribution families assumed to generate each $\bold{x}$ must be justified. Both the multivariate normal and multinomial distributions were chosen because they are sufficiently simple (and well studied) to allow for tractable inference and sufficiently flexible to provide a reasonable approximation to the data gained during extraction. In particular, the multivariate normal distribution over the spatial features $\bold{x}^{s}$ can be thought to represent the shape of each object as an oval; likewise, data generated by each moving object during extraction are often ovular--as noisy extraction procedures cause some smoothing of edges and corners, producing blob-like shapes even when objects are not particularly round--and centered on a given object. Furthermore, this model is justified as the maximum likelihood parameter estimate of a normal distribution corresponds to the least squares fit of data relative to the mean of the distribution. \willie{needs to be said more accurately and explained why this helps.} Modeling the color features $\bold{x}^{c}$ as draws from a multinomial distribution (equivalently, as draws from a product of discrete distributions), is justified since distinct object tend to generate pixels whose hue values are noisy but yield consistent counts in discrete hue bins.

@@

This section details a Markov Chain Monte Carlo (MCMC) sampler used to perform inference, which provides a method to carry out unsupervised detection and tracking. It operates by iteratively sampling from conditional distributions of elements in the joint probability distribution--i.e. it implements Gibbs sampling--in order to gain approximate samples from the posterior distribution of the cluster parameters given the data. Additionally, this sampler makes use of the Metropolis-Hastings (M-H) algorithm to sample from a few distributions.

@@

\begin{algorithm}[h]
\label{alg:MH}
\caption{Metropolis Hastings Algorithm}
\begin{algorithmic}[1]
\STATE $f(x)$ is the desired stationary PDF
\STATE $P(x_{i} | x_{i-1})$ is the proposal pdf
\STATE Initialize $x_{0}$
\FOR{$i = 1 : burn\hspace{1pt}in + N$}
\STATE Draw $x^{*} \sim P(x_{i} | x_{i-1})$
\STATE Draw $\alpha \sim U(0,1)$
\STATE $r_{accept} \leftarrow \frac{f(x^{*})P(x_{i-1} | x^{*})} {f(x_{i-1})P(x^{*} | x_{i-1})} $
\IF{$\alpha < min \{ 1, r_{accept} \} $}
\STATE $x_{i} \leftarrow x^{*}$
\ELSE
\STATE $x_{i} \leftarrow x_{i-1}$
\ENDIF
\ENDFOR
\STATE Discard samples $x_{1}, \ldots, x_{burn\hspace{1pt}in}$
\STATE \textbf{Output}: Remaining samples $x_{1}, \ldots, x_{N}$
\end{algorithmic}
%\caption{Say thing at end}
\end{algorithm}

@@

This section details a Sequential Monte Carlo (SMC) sampler--also known as a particle filter--used to perform inference, which provides another method to carry out unsupervised detection and tracking. An overview of SMC inference is given in Algorithm~\ref{alg:SMC}. \willie{general overview here: at each time step, perform sampling, and then show algorithm figure / thing}

\begin{algorithm}
\caption{Sequential Monte Carlo Inference for the GPUDDPM}
\label{alg:SMC}
\begin{algorithmic}[1]
\FOR{$t = 1 : T$}
\FOR{$l = 1 : L$}
\STATE $K_{t}^{(l)}  \leftarrow K_{t-1}^{(l)}$
\FOR{$i = 1 : N_{t}$}
\STATE Sample assigments from previous time sizes and cluster parameters $\theta_{k,t-1}$
\STATE Draw $c_{i} \sim P(c_{i} | m_{1:K_{t-1}, i-1}, \theta_{1:K_{t-1}, t-1} )$
\IF{$c_{i} = K_{t} + 1$}
\STATE Draw a new cluster $\theta_{K_{t}+1,t} \sim \mathbb{G}_{0}(\text{priors})$   \willie{distro after algo?}
\STATE $K_{t} \leftarrow K_{t} + 1$
\ENDIF
\ENDFOR
\FOR{$s = 1 : S$}
\FOR{$k = 1 : K$}
\STATE Sample new cluster params via aux (given curent assignments and prev clusters)
\ENDFOR
\FOR{$i = 1 : N_{t}$}
\STATE Sample new assignments (given current cluster params)
\STATE Draw $c_{i} \sim P(c_{i} | m_{1:K_{t}, i-1}, \theta_{1:K_{t}, t} )$
\ENDFOR
\STATE Sample new cluster sizes.
\ENDFOR
\ENDFOR
\STATE Only keep some of the $L$ samples.
\ENDFOR
\end{algorithmic}
\end{algorithm}

@@

Noisy extraction procedures may result in erroneous data points that do not correspond to any foreground object. We would like to incorporate these points into our model. To do this, we modify our base distribution $\mathbb{G}_{0}$ to be a two component mixture, where one component is the base distribution and the other is a multivariate normal distribution with a fixed `flat' covarance $\sigma = [100, 0; 0, 100]$ multiplied with an `even' multinomial distribution \willie{need more rigorus terms here} with a fixed parameter $\bold{p} = (1/K, \ldots, 1/K)$.

@@

%, where a numerical search algorith--the Hungarian algorithm \cite{munkres_1957}--is used to find an optimal mapping from ground-truth to output tracks without requiring the performance metric value of all possible mapping combinations to be computed (reducing the time complexity from factorial to polynomial time)

@@

The first synthetic video experiment aimed to test the ability of the model and inference procedure to maintain the identity of independent objects based on color information alone. Two videos were constructed, both containing a red square (rgb value $[255,0,0]$ and size $20$ x $20$ pixels) and a blue square (rgb value $[0,0,255]$ and size $20$ x $20$ pixels). In both videos, the squares begin at opposite sides of the scene at frame $f=1$ and travel towards each other, arriving at the same location at $f=100$ (where the blue square occludes the red square). The second half of the two videos differ in that both squares in the first video continue in the same direction and end at the other's starting positions at $f=200$, and both squares in the second video reverse directions and end at their initial starting positions at $f=200$. Frame difference extraction on each video yields a dataset with identical spatial features; hence, successful tracking depends fully on the reasonable incorporation of color information into the model.

@@

\begin{figure}
  \centering               
  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/cross_final}}
  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/bounce_final}}
  \caption{Each plot shows a sample from the posterior distribution of the state for one of the videos in synthetic experiment one, where the time axis ascends vertically, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown. In both cases, the two objects are successfully tracked through occlusion, though both objects reverse their directions in the second video. \willie{will show assignments by marker type in addition to color}}
  \label{fig:synth_one_plot}
\end{figure}

@@

Benchmark video datasets for object tracking and detection have been produced to provide standard scenes on which researchers can compare detection and tracking results. These videos have been primarily produced for surveillance-related workshops--notably, for the International Workshop on Performance Evaluation of Tracking and Surveillance (PETS)--which provide researchers with video datasets and algorithmic goals on which to focus. Three commonly used benchmark videos from PETS workshops--one used in PETS2000, one in PETS2001, and one used both in PETS2009 and PETS2010--were chosen to demonstrate the efficacy of the method presented in this study. The performance metrics and benchmark datasets allow the methods developed in this paper to be quantitatively compared against other detection and tracking algorithms.

>>>>

% \documentclass[smallcondensed, final, numbook]{svjour3}[12/23/2011]
\documentclass[smallcondensed, final]{svjour3}[12/23/2011]
% \documentclass[twocolumn, final]{svjour3}

@@

\title{Unsupervised Detection and Tracking of Multiple Objects with Dependent Dirichlet Process Mixtures}
\titlerunning{Unsupervised Object Detection and Tracking with Dependent Dirichlet Processes}
% \subtitle{A Bayesian Nonparametric Approach to Tracking Multiple Objects in Videos}
% \subtitle{Subtitle?}
\author{Willie Neiswanger$^{*}$ and Frank Wood$^{\dagger}$}
\institute{
	$^{*}$Columbia University, Department of Applied Math and Applied Physics. 
	Tel.: 503-464-6152.
	\email{wdn2101@columbia.edu}
	\and
	$^{\dagger}$Columbia University, Department of Statistics. 
	Tel.: 212-851-2150. 
	\email{fwood@stat.columbia.edu}}
\date{}
\maketitle
% \mbox{}

\begin{abstract}
This paper proposes an unsupervised technique for the detection and tracking of arbitrary objects in videos. This method is intended to reduce the need for explicit target localization and serve as a general framework applicable to videos with varied object types, backgrounds, and film qualities. The technique uses a dependent Dirichlet process mixture (DDPM) known as the Generalized Polya Urn (GPUDDPM) to model image pixel data that can be extracted in a general manner from the regions in a video that represent objects. This paper describes a specific implementation of the model using spatial and color pixel data extracted via frame differencing and gives two algorithms for performing inference on the model to accomplish detection and tracking. This technique is demonstrated on multiple synthetic and benchmark video datasets that illustrate its ability to, without modification, detect and track variety of objects with diverse physical charactersitics over non-uniform backgrounds and through occlusion.
% \keywords{Object Detection \and Tracking \and Bayesian Nonparametrics \and Multiple Target Tracking}
\end{abstract}

@@

Research in the past ten years involving the development of methods for general detection and tracking of objects has tended to focus on one of either extraction, localization, or tracking; attempts to integrate all three tasks and produce a system aimed at multiple arbitrary objects and diverse video types was not often a primary focus, or was attempted in an ad-hoc manner to test or apply a more specific technique. Research aimed at developing methods to accomplish the three tasks in a cohesive manner has been pursued more recently to produce algorithms that can, in a fully automated way, accomplish unsupervised detection and tracking---or in short, that can determine the spatiotemporal region of a video occupied by each distinct object. A widely applied, though unsophisticated, example of general detection and tracking is the technique of blob tracking. Blob tracking typically refers to methods that perform extraction and localization of objects at each frame independently---for example, by assuming that pixels representing an object are distinguishable from pixels representing the background via color, motion, or texture information, and segmenting the extracted data at each frame into ``blobs'' to accomplish localization. A number of tracking techniques based on finding blobs in consecutive frames of a video with similar locations and appearances have been attempted to maintain the identity of detected blobs over time. The main challenges, however, associated with blob tracking involve the difficulty of segmenting neighboring or occluding objects in a given frame \cite{zhao2004tracking} and preventing multiple tracked blobs from collapsing onto and tracking a single object \cite{vermaak_2003}. These difficulties stem from the typical blob tracker's need to perform detection via segmentation on individual frames and lack of a principled way to perform tracking while resisting the collapse of independent tracks.

This paper presents the use of a Bayesian nonparametric mixture model for general, unsupervised multiple object detection and tracking. This method is intended to provide a widely applicable way to detect arbitrary objects in videos---particularly in cases where frame-by-frame segmentation is difficult or video quality is low and extraction is noisy---and better maintain the isolation of independent objects during tracking. Specifically, this work describes how a Dirichlet process mixture model can accomplish these goals when it is applied to data that may be extracted in a general manner from videos. We will demonstrate how to extract data from objects in videos and how this generative model allows for automated detection and tracking. The model is first presented in an abstract form, to allow it to be used with arbitrary extraction procedures and object appearance representations that may be desired in future studies. We then formulate a specific model for data gained during a basic extraction procedure and describe inference algorithms that allow object localization and tracking to be carried out. We demonstrate our implementation on a number of synthetic and benchmark datasets and describe and compute performance metrics to quantify results.

@@

Methods for tracking single, non-rigid, objects are numerous. Color, for example, has long been used for tracking. These techniques often operate by modeling an object with some color-based appearance model, and using this model to find the object in subsequent frames. One classic approach involves modeling the color of a target (in particular, a distribution over the hue-saturation space of a target region) with a Gaussian mixture model (GMM), and choosing subsequent target positions by searching surrounding areas for regions that yield similar GMMs. Furthermore, the GMM is often allowed to adapt over time, and slowly change to model changes in lighting or other smooth time-based variations in the target's color \citep{raja_1998, mckenna_1999, jepson_2003}. Additionally, color has been successfully applied as the feature in a technique known a the mean-shift procedure (or the kernel-based method) for tracking \citep{comaniciu_2003, perez_2002, nummiaro_2003, lee_2011}. This procedure---a derivation and overview of which can be found in \citep{fukunaga_1975, cheng_1995}--optimizes the search for the region in a frame which is most similar to a target region in the previous frame. It involves an iterative algorithm that repeatedly shifts each data point to the weighted average of data points in its neighborhood; it has been proven that this process converges for each data point. Additionally, the process (in particular, the specification of the neighborhood and the weight-distribution when calculating the weighted mean of nearby data points) is generalized so that multiple kernels can be specified and used to allow for varied clustering behaviors. It can be shown that, through this procedure, each data point becomes associated with a local point of high density (dependent upon the underlying weight distribution specified by the kernel) which naturally allows for clustering \citep{cheng_1995}. The mean shift algorithm has been implemented successfully to allow for a sort of kernel-based object tracking \citep{comaniciu_2003, comaniciu_1999, comaniciu_2000}.

@@

% The mean-shift procedure (also known as `kernel-based object tracking') attempts to provide a robust way for non-rigid objects to be tracked. This method is beneficial because it optimizes the search for the `next' position of an object (i.e. the search for the region in a frame which is most similar to a target region in the previous frame). This procedure---a derivation and overview of which can be found in \citep{fukunaga_1975, cheng_1995}--involves an iterative algorithm that repeatedly shifts each data point to the weighted average of data points in its neighborhood; it has been proven that this process converges for each data point. Additionally, the process (in particular, the specification of the neighborhood and the weight-distribution when calculating the weighted mean of nearby data points) is generalized so that multiple kernels can be specified and used to allow for varied clustering behaviors. It can be shown that, through this procedure, each data point becomes associated with a local point of high density (dependent upon the underlying weight distribution specified by the kernel) which naturally allows for clustering \citep{cheng_1995}. The mean shift algorithm has been implemented successfully to allow for a sort of kernel-based object tracking \citep{comaniciu_2003, comaniciu_1999, comaniciu_2000}. Given the current position of an object at a given frame, the goal of tracking is often to find a nearby position in the next frame that has the most similar distribution over some common set of features. Usually this must be done through an exhaustive search, comparing the similarity of distributions at each nearby position with that at the current position. However, if a certain type of `isotropic kernel' (mean-shift kernel) known as the Bhattacharyya coefficient is chosen as a similarity metric between the feature distributions at two positions, it creates a smooth function where gradient descent techniques can be used to quickly converge upon an optimal subsequent position without using an exhaustive search. Many papers are concerned with applying this kernel-based object tracking scheme to different sets of features or with different kernels. Additionally, the scheme has been attempted with an adaptive kernel whose shape, scale, and orientation is influenced by a target being tracked \citep{wang_2004}, with a kernel adjusted by the estimated centroid of a tracked target \citep{mehmood_2009}, and with a heirarchical version of the mean-shift procedure \citep{dementhon_2002}.

@@

Recall that the distribution over the cluster assignments $c_{i}$ (one for each observation $i$) is a function of the cluster sizes (see \eqref{crp_rep}). Hence, to account for varying numbers of clusters---and in particular, to allow clusters to diminish in size (i.e. reduce the number of assigned observations) and even die---there is a deletion procedure (described in \cite{caron_2007}) by which observations are considered ``removed'' from their assigned clusters at a given time step and value of $m_{k}$ (defined in \eqref{crp_rep} to be the size of cluster $k$) is reduced.

@@

\begin{figure}
  \centering
  \subfloat[]{\label{fig:ants_img}\includegraphics[width=0.335\textwidth]{../img/usbh_frame_723.pdf}} 
  \subfloat[]{\label{fig:ants_img2}\includegraphics[width=0.335\textwidth]{../img/usbh_frame_724.pdf}}
  \subfloat[]{\label{fig:ants_img_framediff}\includegraphics[width=0.323\textwidth]{../img/usbh_framediff.png}}\\
  \subfloat[]{\label{fig:traffic2_img}\includegraphics[width=0.328\textwidth]{../img/traffic2_frame1.png}} 
  \subfloat[]{\label{fig:traffic2_img2}\includegraphics[width=0.328\textwidth]{../img/traffic2_frame2.png}}
  \subfloat[]{\label{fig:traffic2_img_framediff}\includegraphics[width=0.3273\textwidth]{../img/traffic2_framediff.png}}\\
  \subfloat[]{\label{fig:pets2009_img}\includegraphics[width=0.333\textwidth]{../img/pets2009_frame1.png}} 
  \subfloat[]{\label{fig:pets2009_img2}\includegraphics[width=0.333\textwidth]{../img/pets2009_frame2.png}}
  \subfloat[]{\label{fig:pets2009_img_framediff}\includegraphics[width=0.333\textwidth]{../img/pets2009_framediff.png}}\\
  %\subfloat[]{\label{fig:longimg}\includegraphics[width=1.05\textwidth]{../img/longimg_6.pdf}} \\
  \subfloat[]{\label{fig:longimg_all}\includegraphics[width=1\textwidth]{../img/longimg_7_detail.pdf}} 
  \caption{Pairs of consecutive frames and the results produced by taking the pixel-wise difference between these frames. \willie{Want to add colorbar that explains how color represents frame number red = frame 1, blue = frame 100}}
  \label{fig:img_and_framediff}
\end{figure}

@@

The distribution families assumed to generate each $\bold{x}$ must be justified. Both the multivariate normal and multinomial distributions were chosen because they are sufficiently simple (and well studied) to allow for tractable inference and sufficiently flexible to provide a reasonable approximation to the data gained during extraction. In particular, the multivariate normal distribution over the spatial features $\bold{x}^{s}$ can be thought to represent the shape of each object as an oval; likewise, data generated by each moving object during extraction are often ovular---as noisy extraction procedures cause some smoothing of edges and corners, producing blob-like shapes even when objects are not particularly round---and centered on a given object. Furthermore, this model is justified as the maximum likelihood parameter estimate of a normal distribution corresponds to the least squares fit of data relative to the mean of the distribution. \willie{needs to be said more accurately and explained why this helps.} Modeling the color features $\bold{x}^{c}$ as draws from a multinomial distribution (equivalently, as draws from a product of discrete distributions), is justified since distinct object tend to generate pixels whose hue values are noisy but yield consistent counts in discrete hue bins.

@@

This section details a Markov Chain Monte Carlo (MCMC) sampler used to perform inference, which provides a method to carry out unsupervised detection and tracking. It operates by iteratively sampling from conditional distributions of elements in the joint probability distribution---i.e. it implements Gibbs sampling---in order to gain approximate samples from the posterior distribution of the cluster parameters given the data. Additionally, this sampler makes use of the Metropolis-Hastings (M-H) algorithm to sample from a few distributions.

@@

\begin{algorithm}[h!]
\label{alg:MH}
\caption{Metropolis Hastings Algorithm}
\begin{algorithmic}[1]
\STATE $f(x)$ is the desired stationary PDF
\STATE $P(x_{i} | x_{i-1})$ is the proposal pdf
\STATE Initialize $x_{0}$
\FOR{$i = 1 : burn\hspace{1pt}in + N$}
\STATE Draw $x^{*} \sim P(x_{i} | x_{i-1})$
\STATE Draw $\alpha \sim U(0,1)$
\STATE $r_{accept} \leftarrow \frac{f(x^{*})P(x_{i-1} | x^{*})} {f(x_{i-1})P(x^{*} | x_{i-1})} $
\IF{$\alpha < min \{ 1, r_{accept} \} $}
\STATE $x_{i} \leftarrow x^{*}$
\ELSE
\STATE $x_{i} \leftarrow x_{i-1}$
\ENDIF
\ENDFOR
\STATE Discard samples $x_{1}, \ldots, x_{burn\hspace{1pt}in}$
\STATE \textbf{Output}: Remaining samples $x_{1}, \ldots, x_{N}$
\end{algorithmic}
%\caption{Say thing at end}
\end{algorithm}

@@

This section details a Sequential Monte Carlo (SMC) sampler---also known as a particle filter---used to perform inference, which provides another method to carry out unsupervised detection and tracking. The SMC inference algorithm is given in Algorithm~\ref{alg:SMC}. The algorithm operates by

\begin{algorithm}
\caption{Sequential Monte Carlo Inference for the GPUDDPM}
\label{alg:SMC}
\begin{algorithmic}[1]
\FOR{$t = 1 : T$}
\FOR{$l = 1 : L$}
\STATE $K_{t}^{(l)}  \leftarrow K_{t-1}^{(l)}$
\FOR{$i = 1 : N_{t}$}
\STATE Sample assigments from previous time sizes and cluster parameters $\theta_{k,t-1}$
\STATE Draw $c_{i} \sim P(c_{i} | m_{1:K_{t-1}, i-1}, \theta_{1:K_{t-1}, t-1} )$
\IF{$c_{i} = K_{t} + 1$}
\STATE Draw a new cluster $\theta_{K_{t}+1,t} \sim \mathbb{G}_{0}(\text{priors})$   \willie{distro after algo?}
\STATE $K_{t} \leftarrow K_{t} + 1$
\ENDIF
\ENDFOR
\FOR{$s = 1 : S$}
\FOR{$k = 1 : K$}
\STATE Sample new cluster params via aux (given curent assignments and prev clusters)
\ENDFOR
\FOR{$i = 1 : N_{t}$}
\STATE Sample new assignments (given current cluster params)
\STATE Draw $c_{i} \sim P(c_{i} | m_{1:K_{t}, i-1}, \theta_{1:K_{t}, t} )$
\IF{$c_{i} = K_{t} + 1$}
\STATE Draw a new cluster $\theta_{K_{t}+1,t} \sim \mathbb{G}_{0}(\text{priors})$   \willie{distro after algo?}
\STATE $K_{t} \leftarrow K_{t} + 1$
\ENDIF
\ENDFOR
\STATE Sample new cluster sizes.
\ENDFOR
\ENDFOR
\STATE Only keep some of the $L$ samples.
\ENDFOR
\end{algorithmic}
\end{algorithm}

A few key distributions used in the SMC algorithm are shown below.

@@

Noisy extraction procedures may result in spurrious data points that do not correspond to any foreground object. We would like to incorporate these points into our model. To do this, we modify our base distribution $\mathbb{G}_{0}$ to be a two component mixture, where one component is the base distribution and the other is a multivariate normal distribution with a fixed `flat' covarance $\sigma = [100, 0; 0, 100]$ multiplied with an `even' multinomial distribution \willie{need more rigorus terms here} with a fixed parameter $\bold{p} = (1/K, \ldots, 1/K)$.

@@

%, where a numerical search algorith---the Hungarian algorithm \cite{munkres_1957}--is used to find an optimal mapping from ground-truth to output tracks without requiring the performance metric value of all possible mapping combinations to be computed (reducing the time complexity from factorial to polynomial time)

@@

The first synthetic video experiment aimed to test the ability of the model and inference procedure to maintain the identity of independent objects based on color information alone. Two videos were constructed, both containing a red square (rgb value $[255,0,0]$ and size $20$ x $20$ pixels) and a blue square (rgb value $[0,0,255]$ and size $20$ x $20$ pixels). In both videos, the squares begin at opposite sides of the scene at frame $f=1$ and travel towards each other, arriving at the same location at $f=100$ (where the blue square occludes the red square). The second half of the two videos differ in that both squares in the first video continue in the same direction and end at the other's starting positions at $f=200$, and both squares in the second video reverse directions and end at their initial starting positions at $f=200$. The frame difference extraction yields identical spatial features in both videos; hence, successful tracking depends fully on the incorporation of color information into the model.

@@

Groundtruth for the synthetic videos

\begin{figure}
  \centering               
  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/synth1_bounce.pdf}}
  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/synth1_cross.pdf}}
  \caption{Each plot shows a sample from the posterior distribution of the state for one of the videos in synthetic experiment one, where the time axis ascends vertically, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown. In both cases, the two objects are successfully tracked through occlusion, though both objects reverse their directions in the second video. \willie{will show assignments by marker type in addition to color}}
  \label{fig:synth_one_plot}
\end{figure}

@@

Benchmark video datasets for object tracking and detection have been produced to provide standard scenes on which researchers can compare detection and tracking results. These videos have been primarily produced for surveillance-related workshops---notably, for the International Workshop on Performance Evaluation of Tracking and Surveillance (PETS)--which provide researchers with video datasets and algorithmic goals on which to focus. Three commonly used benchmark videos from PETS workshops---one used in PETS2000, one in PETS2001, and one used both in PETS2009 and PETS2010---were chosen to demonstrate the efficacy of the method presented in this study. The performance metrics and benchmark datasets allow the methods developed in this paper to be quantitatively compared against other detection and tracking algorithms.
